Carbonate chemistry at Ny-Ålesund

### 2023-03-10 awipev-CO2.Rmd
More fig refining

# for paper
at_s <- d_all_at_discrete %>% 
  dplyr::select(sal_fb_mix, at) %>% 
  drop_na() %>% 
  ggplot(aes(x = sal_fb_mix, y = at)) +
  geom_point(col="blue", size = 1) +
  labs(x= latex2exp::TeX("Salinity"),
       y= latex2exp::TeX("Total alkalinity ($\\mu mol\\phantom{x}kg^{-1})$"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:OLS", se = FALSE, size = 1) +
  #ggpmisc::stat_ma_eq(method = "lmodel2:OLS", aes(label = after_stat(eq.label))) +
  theme_bw() + Mytheme(size_labs = 12) + theme(aspect.ratio=1) + 
  coord_fixed()
ggsave(filename = "figures/at_s.png", plot = at_s, width = 25, height = 25, units = "cm")

@@

# for paper
pco2_calc_meas <- d_all %>% 
  dplyr::select(pco2_fb_insi_temp, pco2_calc_fb_insi_temp) %>% 
  drop_na() %>% 
  ggplot(aes(x = pco2_fb_insi_temp, y = pco2_calc_fb_insi_temp)) +
  geom_point(col="blue", size = 1) +
  xlim(180, 420) + ylim(180, 420) +
  labs(x=latex2exp::TeX("Measured $pCO_{2}\\phantom{x}(\\mu atm)$"),
       y= latex2exp::TeX("Calculated $pCO_{2}\\phantom{x}(\\mu atm)$"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  #ggpmisc::stat_ma_eq(method = "lmodel2:MA", aes(label = after_stat(eq.label))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 12) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
ggsave(filename = "figures/pco2_calc_meas.png", plot = pco2_calc_meas, width = 25, height = 25, units = "cm")
```

@@

offset_ph <- ph_sf %>% 
  drop_na() %>% 
  ggplot() +
  geom_point(aes(x=datetime, y=diff_spec_sf), size=1, color="blue") +
  geom_hline(yintercept=0, size=0.3) +
  geom_hline(yintercept=-0.02, size=0.3, linetype = "dashed") +
  geom_hline(yintercept=0.02, size=0.3, linetype = "dashed") +
  labs(x = "Date", y=expression(paste(pH[T], " difference")), title = NULL) +
  Mytheme(size_labs = 12, face_font = "plain") + theme(aspect.ratio=1)  + 
  scale_colour_discrete(guide="none") + 
  coord_fixed()
ggsave(filename = "figures/offset_ph.png", plot = offset_ph, width = 25, height = 25, units = "cm")
```

@@

pHspec_pHseafet <- d_all %>% 
  dplyr::select(ph_s_sf_temp_insi, pHint_tot_sf) %>% 
  drop_na() %>% 
  ggplot(aes(x = ph_s_sf_temp_insi, y = pHint_tot_sf)) +
  geom_point(col="blue", size = 1) +
  xlim(8.05, 8.18) + ylim(8.05, 8.18) +
  labs(x="Spectrophotometric pH",
       y= "SeaFET pH")  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  #ggpmisc::stat_ma_eq(method = "lmodel2:MA", aes(label = after_stat(eq.label), digits = 4)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 12) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
ggsave(filename = "figures/pHspec_pHseafet.png", plot = pHspec_pHseafet, width = 25, height = 25, units = "cm")
```

@@

ts_sal <- d_all %>%
  ggplot(aes(x = datetime, y = sal_fb_mix), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title=NULL,x="", y="Salinity") +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs))

# temp_insitu_11m
ts_temp <- d_all %>%
  ggplot(aes(x = datetime, y = temp_insitu_11m), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title=NULL,x="", y="Temperature (°C)") +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs))

# PAR
ts_par_air <- d_all %>%
  ggplot(aes(x = datetime, y = par_air), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title="Atmospheric PAR",x="", 
       y=latex2exp::TeX("PAR ($\\mu mol\\phantom{x}m^{-2}\\phantom{x}s^{-1})$")) +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs))

# PCO2 
ts_pCO2 <- d_all %>%
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pco2_fb_insi_temp), 
             colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  geom_point(aes(x = datetime, y = pco2_calc_fb_insi_temp), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title=NULL,
       subtitle = NULL,
       x="", y=latex2exp::TeX("pCO$_2$ ($\\mu atm$)", bold=TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs))

# SEAFET pH
ts_pH <- d_all %>% 
  dplyr::filter(depth > 8) %>% 
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pHint_tot_sf), 
             colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) + 
  geom_point(aes(x = datetime, y = ph_s_sf_temp_insi), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title=NULL, 
       subtitle =NULL,
       x = "", y=latex2exp::TeX("pH$_T$", bold=TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(legend.position = c(0.1, 0.5),
        legend.key.height = unit(0.5, 'cm')) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs))

# AT
ts_at <- d_all %>%
  ggplot(aes(x = datetime, y = at_calc), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  geom_point(aes(x = datetime, y = at), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title=NULL, 
       subtitle = NULL,
       x="", y=latex2exp::TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$", bold = TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs))

# Combine plots
g1 <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at)
g1 <- g1 + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(face = "bold"))
ggsave(filename = "figures/time_series.png", plot = g1, width = 15, height = 30, units = "cm")

@@

  labs(title="", y = "Salinity", x=NULL) +

@@

        aspect.ratio=0.5,

@@

        axis.title.y = element_text(face = "bold", size = size_labs)

@@

  labs(title=NULL, y = "Temperature (°C)", x=NULL) +

@@

  labs(title=NULL, y = latex2exp::TeX("pCO$_2$ ($\\mu atm$)", bold = TRUE), x=NULL) +

@@

  labs(title=NULL, y = latex2exp::TeX("$pH_T$", bold = TRUE), x=NULL) +

@@

  labs(title=NULL, y = TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$", bold = TRUE), x=NULL) +

@@

ggsave(filename = "figures/ts_boxplots.png", plot = g2)

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
g <- (at_s | pco2_calc_meas) / (offset_ph | pHspec_pHseafet)
g <- g + patchwork::plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(face = "bold"))
ggsave(filename = "figures/pco2_ph_etc.png", plot = g)
ggsave(filename = "figures/essd/pco2_ph_etc.pdf", plot = g)
```

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
g <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at) |
  (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g <- g + patchwork::plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(face = "bold", size = 14))
ggsave(filename = "figures/essd/ts.pdf", plot = g, width = 20, height = 30)
# legend:
# - Salinity below 8m
# - In situ temperature (11m)
# - pCO2 at 11 m and insitu Temperature, measued in blue, calculated from CT an AT in red
# - In situ pH below 8 m at in situ Temperature, seaFET in blue, spectrophotometric in red
# - Total alkalinity calculated from salinity in blue, discrete measurement in red

>>>>

# for paper
at_s <- d_all_at_discrete %>% 
  dplyr::select(sal_fb_mix, at) %>% 
  drop_na() %>% 
  ggplot(aes(x = sal_fb_mix, y = at)) +
  geom_point(col="blue", size = 1) +
  labs(title = "A",
       x= latex2exp::TeX("\\textbf{Salinity}"),
       y= latex2exp::TeX("\\textbf{Total alkalinity ($\\mu mol\\phantom{x}kg^{-1})$}"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:OLS", se = FALSE, size = 1) +
  #ggpmisc::stat_ma_eq(method = "lmodel2:OLS", aes(label = after_stat(eq.label))) +
  theme_bw() + 
  Mytheme(size_labs = 12) +
  theme(aspect.ratio=1, plot.title = element_text(face = "bold"),
        plot.margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")) + 
  coord_fixed()
ggsave(filename = "figures/at_s.png", plot = at_s, width = 25, height = 25, units = "cm")

@@

# for paper
pco2_calc_meas <- d_all %>% 
  dplyr::select(pco2_fb_insi_temp, pco2_calc_fb_insi_temp) %>% 
  drop_na() %>% 
  ggplot(aes(x = pco2_fb_insi_temp, y = pco2_calc_fb_insi_temp)) +
  geom_point(col="blue", size = 1) +
  xlim(180, 420) + ylim(180, 420) +
  labs(title = "B",
       x=latex2exp::TeX("\\textbf{Measured $pCO_{2}\\phantom{x}(\\mu atm)$}"),
       y= latex2exp::TeX("\\textbf{Calculated $pCO_{2}\\phantom{x}(\\mu atm)$}"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  #ggpmisc::stat_ma_eq(method = "lmodel2:MA", aes(label = after_stat(eq.label))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 12) + 
  Mytheme(size_labs = 12) +
  theme(aspect.ratio=1, plot.title = element_text(face = "bold"),
        plot.margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")) + 
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
ggsave(filename = "figures/pco2_calc_meas.png", plot = pco2_calc_meas, width = 25, height = 25, units = "cm")
```

@@

offset_ph <- ph_sf %>% 
  drop_na() %>% 
  ggplot() +
  geom_point(aes(x=datetime, y=diff_spec_sf), size=1, color="blue") +
  geom_hline(yintercept=0, size=0.3) +
  geom_hline(yintercept=-0.02, size=0.3, linetype = "dashed") +
  geom_hline(yintercept=0.02, size=0.3, linetype = "dashed") +
  labs(title = "C",
       x = "Date", y=latex2exp::TeX("\\textbf{pH$_T$ difference}")) +
  #Mytheme(size_labs = 12, face_font = "plain") + 
  Mytheme(size_labs = 12) +
  theme(aspect.ratio=1, plot.title = element_text(face = "bold"),
        plot.margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold")) + 
  scale_colour_discrete(guide="none") + 
  coord_fixed()
ggsave(filename = "figures/offset_ph.png", plot = offset_ph, width = 25, height = 25, units = "cm")
```

@@

pHspec_pHseafet <- d_all %>% 
  dplyr::select(ph_s_sf_temp_insi, pHint_tot_sf) %>% 
  drop_na() %>% 
  ggplot(aes(x = ph_s_sf_temp_insi, y = pHint_tot_sf)) +
  geom_point(col="blue", size = 1) +
  xlim(8.05, 8.18) + ylim(8.05, 8.18) +
  labs(title = "D",
       x="Spectrophotometric pH",
       y= "SeaFET pH")  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  #ggpmisc::stat_ma_eq(method = "lmodel2:MA", aes(label = after_stat(eq.label), digits = 4)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + 
  Mytheme(size_labs = 12) +
  theme(aspect.ratio=1, plot.title = element_text(face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        plot.margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")) + 
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
ggsave(filename = "figures/pHspec_pHseafet.png", plot = pHspec_pHseafet, width = 25, height = 25, units = "cm")
```

@@

ts_sal <- d_all %>%
  ggplot(aes(x = datetime, y = sal_fb_mix), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title="A",x="", y="Salinity") +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs),
        title = element_text(colour="black", size=14, face="bold")
        )

# temp_insitu_11m
ts_temp <- d_all %>%
  ggplot(aes(x = datetime, y = temp_insitu_11m), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title="B",x="", y="Temperature (°C)") +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs),
        title = element_text(colour="black", size=14, face="bold")
        )
# PAR
ts_par_air <- d_all %>%
  ggplot(aes(x = datetime, y = par_air), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title="Atmospheric PAR",x="", 
       y=latex2exp::TeX("PAR ($\\mu mol\\phantom{x}m^{-2}\\phantom{x}s^{-1})$")) +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs))

# PCO2 
ts_pCO2 <- d_all %>%
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pco2_fb_insi_temp), 
             colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  geom_point(aes(x = datetime, y = pco2_calc_fb_insi_temp), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title="C",
       subtitle = NULL,
       x="", y=latex2exp::TeX("pCO$_2$ ($\\mu atm$)", bold=TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs),
        title = element_text(colour="black", size=14, face="bold")
        )
# SEAFET pH
ts_pH <- d_all %>% 
  dplyr::filter(depth > 8) %>% 
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pHint_tot_sf), 
             colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) + 
  geom_point(aes(x = datetime, y = ph_s_sf_temp_insi), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title="D", 
       subtitle =NULL,
       x = "", y=latex2exp::TeX("pH$_T$", bold=TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(legend.position = c(0.1, 0.5),
        legend.key.height = unit(0.5, 'cm')) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs),
        title = element_text(colour="black", size=14, face="bold")
        )
# AT
ts_at <- d_all %>%
  ggplot(aes(x = datetime, y = at_calc), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  geom_point(aes(x = datetime, y = at), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title="E", 
       subtitle = NULL,
       x="", y=latex2exp::TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$", bold = TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs),
        title = element_text(colour="black", size=14, face="bold")
        )

# Combine plots
g1 <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at)
#g1 <- g1 + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(face = "bold"))
ggsave(filename = "figures/time_series.png", plot = g1, width = 15, height = 30, units = "cm")

@@

  labs(title="F", y = "Salinity", x=NULL) +

@@

#        aspect.ratio=0.5,

@@

        axis.title.y = element_text(face = "bold", size = size_labs),
        title = element_text(colour="black", size=14, face="bold")

@@

  labs(title="G", y = "Temperature (°C)", x=NULL) +

@@

  labs(title="H", y = latex2exp::TeX("pCO$_2$ ($\\mu atm$)", bold = TRUE), x=NULL) +

@@

  labs(title="I", y = latex2exp::TeX("$pH_T$", bold = TRUE), x=NULL) +

@@

  labs(title="J", y = TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$", bold = TRUE), x=NULL) +

@@

ggsave(filename = "figures/ts_boxplots.png", plot = g2, width = 12)

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
g <- (at_s | pco2_calc_meas) / (offset_ph | pHspec_pHseafet) #+
  #plot_layout(heights = c(6,6,6,6)) 
# g <- g + patchwork::plot_annotation(tag_levels = 'A') & 
#   theme(plot.tag = element_text(face = "bold"))
ggsave(filename = "figures/pco2_ph_etc.png", plot = g)
ggsave(filename = "figures/essd/pco2_ph_etc.pdf", plot = g, width = 18, height = 18, units = "cm")
```

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
g <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at) |
  (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
#g <- g + patchwork::plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(face = "bold", size = 14))
ggsave(filename = "figures/essd/ts.pdf", plot = g, width = 12)
# legend:
# - Salinity below 8m
# - In situ temperature (11m)
# - pCO2 at 11 m and insitu Temperature, measued in blue, calculated from CT an AT in red
# - In situ pH below 8 m at in situ Temperature, seaFET in blue, spectrophotometric in red
# - Total alkalinity calculated from salinity in blue, discrete measurement in red
